---
title: "Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

After each class, work in groups of 2 or 3 to produce an RMarkdown document in which you answer the questions below. Knit your document into a `.pdf` file and submit it to the Moodle discussion forum for feedback. For each question, explain and justify your approach and conclude briefly.

Here is an example of header for your RMarkdown document:

```
---
title: "Homework #1"
author: "Group: the kurtotic avengers"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---
```

For the pdf output to work, you need TeX installed on your computer. On a Mac, try [MacTeX](http://www.tug.org/mactex/), on a PC, see options [here](https://www.latex-project.org/get/). If you cannot get a pdf output, try exporting as Word or html (change `pdf_document` to `word_document` or `html_document` in the header), then print the html document as pdf.

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 1</button>

## Exercise 1

Illustrate the results from the dice exercise using the code in the [activities](activities.html) tab.

Apply at least two inferential statistical tests of your choice to each pairwise combinations of groups.

- Do you have sufficient evidence to suggest that the groups were sampled from different populations?  

- In what way do you think the groups differ?  

## Exercise 2

Compute one-sample bootstrap confidence intervals of the mean for each group.

Compute bootstrap confidence intervals for the difference in means between two groups of the dice data.

Do that for all pairwise combinations.

- Describe the results using numerical values, graphical representations, or both.

- Do you have sufficient evidence to suggest that the groups were sampled from different populations?

## Exercise 3

Recently I've started to write down by how much my morning bus is delayed. 
You can load the data in the next chunk. 

Each value is a delay relative to the official bus timetable:

- a positive value means the bus was late;

- a negative value means the bus was early;

- zero means the bus was exactly on time, within 60 seconds (we're not in Japan, so I only check the minutes on my watch when the bus arrives, not the seconds).

```{r}
delays <- matrix(scan("./data/bustimes.txt"), byrow=TRUE) # delays in minutes
```

- [1] Illustrate the results.
- [2] What is the typical waiting time? Provide a confidence interval. You can try several methods.
- [3] Let say I have the hypothesis that I tend to wait at least 3 minutes for my bus in the morning. Given the data, and using bootstrap inferences, what can you conclude about this hypothesis?

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 2</button>

## Exercise 1

- Repeat exercise 1 from last class. 

- How do the current results differ relative to the results from last class?

- Repeat the group comparisons after combining data across classes, separately for each colour.

- For each dataset, check for outliers using at least one robust method.


## Exercise 2

- Simulate the sampling distribution of 10,000 throws of a dice with 6 facets.

- Simulate the same sampling distribution, but for a biased dice, which is twice as likely to land on 6 than on any other facet. 

- In a figure, superimpose the two distributions. How do they differ?

Hint: use the `sample()` function. The bias can be controlled using the `prob` argument.

## Exercise 3

Consider a standard lognormal population. Compare the sampling distributions for the mean, the median and a trimmed mean for sample sizes n = 10, 20, 50, 100. Use 10,000 simulations to derive the sampling distributions.
Illustrate the sampling distributions using ggplot2. Try using `facet_grid` or `facet_wrap` to combine multiple plots into one figure. Finally, report the standard error for each estimator and sample size. What do you conclude?

### Generate a random sample of size n=10 from a standard lognormal population
```{r, eval=FALSE}
set.seed(21) # reproducible example
n <- 10
rlnorm(n)
```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 3</button>

## Exercise 1

- Repeat exercise 1 from last class. 

- How do the current results differ relative to the results from last class?

- Repeat the group comparisons after combining data across classes, separately for each colour.

## Exercise 2
```{r, eval=FALSE}

```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 4</button>

## Exercise 1

- Repeat exercise 1 from last class. 

- How do the current results differ relative to the results from last class?

- Repeat the group comparisons after combining data across classes, separately for each colour.

## Exercise 2
```{r, eval=FALSE}

```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 5</button>

## Exercise 1
```{r, eval=FALSE}

```

## Exercise 2
```{r, eval=FALSE}

```




