---
title: "Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

After each class, work in groups of 2 or 3 to produce an RMarkdown document in which you answer the questions below. Knit your document into a `.pdf` file and submit it to the Slack channel for feeback.

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 1</button>

## Question 1
```{r, eval=FALSE}

```

## Question 2

Compute bootstrap confidence intervals for the difference between two groups of the dice data.

```{r, eval=FALSE}

```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 2</button>

## Question 1
```{r, eval=FALSE}

```

## Question 2
```{r, eval=FALSE}

```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 3</button>

## Question 1
```{r, eval=FALSE}

```

## Question 2
```{r, eval=FALSE}

```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 4</button>

## Question 1
```{r, eval=FALSE}

```

## Question 2
```{r, eval=FALSE}

```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 5</button>

## Question 1
```{r, eval=FALSE}

```

## Question 2
```{r, eval=FALSE}

```




