---
title: "Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

After each class, work in groups of 2 or 3 to produce an RMarkdown document in which you answer the questions below. Knit your document into a `.pdf` file and submit it to the Moodle discussion forum for feedback. For each question, explain and justify your approach and conclude briefly.

Here is an example of header for your RMarkdown document:

```
---
title: "Homework #1"
author: "Group: the kurtotic avengers"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---
```

For the pdf output to work, you need TeX installed on your computer. On a Mac, try [MacTeX](http://www.tug.org/mactex/), on a PC, see options [here](https://www.latex-project.org/get/). If you cannot get a pdf output, try exporting as Word or html (change `pdf_document` to `word_document` or `html_document` in the header), then print the html document as pdf.

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 1</button>

## Exercise 1

Illustrate the results from the dice exercise using the code in the [activities](activities.html) tab.

Apply at least two inferential statistical tests of your choice to each pairwise combinations of groups.

- Do you have sufficient evidence to suggest that the groups were sampled from different populations?  

- In what way do you think the groups differ?  

## Exercise 2

Compute one-sample bootstrap confidence intervals of the mean for each group.

Compute bootstrap confidence intervals for the difference in means between two groups of the dice data.

Do that for all pairwise combinations.

- Describe the results using numerical values, graphical representations, or both.

- Do you have sufficient evidence to suggest that the groups were sampled from different populations?

## Exercise 3

Recently I've started to write down by how much my morning bus is delayed. 
You can load the data in the next chunk. 

Each value is a delay relative to the official bus timetable:

- a positive value means the bus was late;

- a negative value means the bus was early;

- zero means the bus was exactly on time, within 60 seconds (we're not in Japan, so I only check the minutes on my watch when the bus arrives, not the seconds).

```{r}
delays <- matrix(scan("./data/bustimes.txt"), byrow=TRUE) # delays in minutes
```

- [1] Illustrate the results.
- [2] What is the typical waiting time? Provide a confidence interval. You can try several methods.
- [3] Let say I have the hypothesis that I tend to wait at least 3 minutes for my bus in the morning. Given the data, and using bootstrap inferences, what can you conclude about this hypothesis?

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 2</button>

## Exercise 1

- Repeat exercise 1 from last class. 

- How do the current results differ relative to the results from last class?

- Repeat the group comparisons after combining data across classes, separately for each colour.

- For each dataset, check for outliers using at least one robust method.


## Exercise 2

- Simulate the sampling distribution of 10,000 throws of a dice with 6 facets.

- Simulate the same sampling distribution, but for a biased dice, which is twice as likely to land on 6 than on any other facet. 

- In a figure, superimpose the two distributions. How do they differ?

Hint: use the `sample()` function. The bias can be controlled using the `prob` argument.

## Exercise 3

Consider a standard lognormal population. Compare the sampling distributions for the mean, the median and a trimmed mean for sample sizes n = 10, 20, 50, 100. Use 10,000 simulations to derive the sampling distributions.
Illustrate the sampling distributions using ggplot2. Try using `facet_grid` or `facet_wrap` to combine multiple plots into one figure. Finally, report the standard error for each estimator and sample size. What do you conclude?

### Generate a random sample of size n=10 from a standard lognormal population
```{r, eval=FALSE}
set.seed(21) # reproducible example
n <- 10
rlnorm(n)
```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 3</button>

## Exercise 1

- Illustrate the dice results from the different classes in one figure with multiple panels.

- How do the results differ among classes? What do you conclude?

- Perform group comparisons after combining data across classes.

- Use a selection of approaches to compute CIs using the mean and a trimmed mean in combination with a standard parametric method, the percentile bootstrap and the bootstrap-t. Do not report CIs for all combinations of choices: instead justify carefully your choice of one or two methods based on what we saw in class and the assigned references. By now you know that using standard t-tests on means should not be the default because there are several potentially more appropriate alternatives.

## Exercise 2

The next chunk contains data from Germany, in different years, of the number of brooding pairs of storks and the millions of births of human babies.

- Illustrate the results.
- Is there a relationship between the stork population and human births?
- Compute Pearson and Spearman correlation coefficients.
- Compute CIs using standard methods (implemented in `cor.test()`) and the percentile bootstrap.
- What do you conclude?

```{r, eval=FALSE}
years  <- c(1965, 1971, 1974, 1977, 1978, 1979, 1980)
storks <- c(1920, 1700, 1090, 990, 1030,  995,  930)
babies <- c(1.04, 0.78, 0.62, 0.58, 0.57, 0.58, 0.62)
```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 4</button>

## Exercise 1

- Illustrate the dice results from the different classes in one figure with multiple panels.

- Based on the graphical representations (do not use inferential tests), how do the results differ among classes? What do you conclude?

- Perform group comparisons only after combining data across the 4 classes. Choose carefully 2 approaches based on everything you have learnt so far. Justify your choices, embed the results in text format using [inline code](https://rmarkdown.rstudio.com/lesson-4.html), and conclude succinctly but carefully.

## Exercise 2

Here are data from a weight gain programme. Weights in pounds were measured before (`t1`) and after (`t2`) the programme in the same participants (`t1[1]` and `t2[1]` are measurements from the same person). 

- Illustrate the results.

- Using inferential statistics and graphical representations, do you have sufficient evidence to conclude that the programme is effective?

- If you had the task to promote the programme, what would be the typical weight change you would advertise?

```{r, eval=FALSE}
t1 <- c(83.8, 83.3, 86, 82.5, 86.7, 79.6, 76.9, 94.2, 73.4, 80.5, 81.6, 82.1, 77.6, 83.5, 89.9, 86, 87.3)
t2 <- c(95.2, 94.3, 91.5, 91.9, 100.3, 76.7, 76.8, 101.6, 94.9, 75.2, 77.8, 95.5, 90.7, 92.5, 93.8, 91.7, 98)
```

<button type="button" class="btn btn-success btn-lg btn-block font-weight-bold">Class 5</button>

## Exercise 1

[Ritchie et al. (2018)](https://academic.oup.com/cercor/article/28/8/2959/4996558) reported sex differences in the volume of adult human brains from 5216 UK Biobank participants. They found larger mean brain volumes in men than in women (look at the supplementary information to see how they used some of the shift function code to illllustrate the effects!). They also found increased variance in the sample of men relative to the sample of women. The means and standard deviations of total brain volumes for men and women reported by Ritchie et al. are available in the next chunk. 

```{r, eval=FALSE}
w.m <- 1115.76
m.m <- 1233.58 
w.sd <- 89.68
m.sd <- 98.31
```

You plan a new experiment in which you would like to replicate the variance difference in total brain volume. Assuming the results from Ritchie et al. are approximately normally distributed, use the values from the previous chunk to simulate experiments to determine the minimum sample size to achieve at least 90% statistical power.

To compare variances, a modified percentile bootstrap test is implemented in `comvar2()`. The syntax is `comvar2(x,y,nboot=1000)`. The function return several outputs, including a confidence interval, but no *P* value. You can source the stand-alone function using the next chunk:

```{r}
source('./code/comvar2.txt')
```

Illustrate the results and conclude.

## Exercise 2

Repeat exercise 1 using the Kolmogorov-Smirnov test. Illustrate the results and conclude.

## Exercise 3

Simulate two large normal samples of *n*=1000 each using the summary statistics from exercise 1. Compare the two samples using a shift function. Conclude.




