---
title: "Notebook 1: Introduction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies
```{r}
library(tibble)
library(ggplot2)
```

# Rand Wilcox's code
When I mention code from Rand Wilcox, it is available in a giant text file.
You access the functions by using the `source()` function:
```{r, eval=FALSE}
source("./code/Rallfun-v35.txt")
```

In some of the notebooks, I've extracted one or a few functions so you can source a smaller file. This prevents your environment from being cluttered with functions you don't need. Creating a well documented R package is a lot of work, so some statisticians don't spend the time required. But that means their code is less accessible.

See more detail and updated versions of the code [here](https://dornsife.usc.edu/labs/rwilcox/software/).

Some of the functions are also available in the `WRS2` package.

# Dice exercise

## Data capture

Enter the data in R or a text file: 5 rows (trials) and 10 columns (participants).

```{r}
# enter data in matrix with 5 rows, 10 columns:

# dice_res[,1] <- c(,,,,)
# dice_res[,2] <- c(,,,,)
# dice_res[,3] <- c(,,,,)
# dice_res[,4] <- c(,,,,)
# dice_res[,5] <- c(,,,,)
# dice_res[,6] <- c(,,,,)
# dice_res[,7] <- c(,,,,)
# dice_res[,8] <- c(,,,,)
# dice_res[,9] <- c(,,,,)
# dice_res[,10] <- c(,,,,)

# Save in a txt file:

# write.table(dice_res, file = "./data/dice_res_blue.txt", row.names=FALSE, col.names=FALSE)
# write.table(dice_res, file = "./data/dice_res_green.txt", row.names=FALSE, col.names=FALSE)
# write.table(dice_res, file = "./data/dice_res_red.txt", row.names=FALSE, col.names=FALSE)
```

## Load simulated data

```{r}
nt <- 5
np <- 10
res_red <- matrix(scan("./data/res1.txt"), nrow=nt, byrow=TRUE)
res_green <- matrix(scan("./data/res2.txt"), nrow=nt, byrow=TRUE)
res_blue <- matrix(scan("./data/res3.txt"), nrow=nt, byrow=TRUE)
```

## Load real data

```{r, eval = FALSE}
res_red <- matrix(scan("./data/dice_res_red.txt"), nrow=nt, byrow=TRUE)
res_green <- matrix(scan("./data/dice_res_green.txt"), nrow=nt, byrow=TRUE)
res_blue <- matrix(scan("./data/dice_res_blue.txt"), nrow=nt, byrow=TRUE)
```

## Illustration

Mean results
```{r}
m.red <- round(apply(res_red, 2, mean), digits=1)
m.green <- round(apply(res_green, 2, mean), digits=1)
m.blue <- round(apply(res_blue, 2, mean), digits=1)
```

Illustrate results
```{r}
set.seed(1)
df <- tibble(obs = c(m.red, m.green, m.blue),
             gp = factor(c(rep("Red",np),rep("Green",np),rep("Blue",np)))
             )

# preserve factor order
df$gp <- as.character(df$gp)
df$gp <- factor(df$gp, levels=unique(df$gp))

  ggplot(df, aes(gp, obs)) + theme_linedraw() +
  geom_point(aes(colour = gp), position = position_jitter(width = 0.1)) +
    scale_color_manual(values=c("red", "green4", "blue")) +
  theme(legend.position = "none",
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16)) +
  labs(x = "Groups", y = "Observations") +   
    coord_cartesian(ylim = c(0, 12)) +
    scale_y_continuous(breaks = seq(0, 12, 2)) +
  geom_errorbar(stat = "summary", fun.y = mean, fun.ymin = mean,
                fun.ymax = mean, colour = "grey", width = .55, size = 3) +
  geom_errorbar(stat = "summary", fun.y = mean, fun.ymin = mean,
                fun.ymax = mean, colour = "black", width = .5, size = 1)
```

## Inferential statistics?

What would you do next with the data?

# 50 experiments - 50 confidence intervals

We perform many experiments, and for each experiment we compute a confidence interval, which we plot as horizontal lines.

## Sample from a normal distribution
```{r}
nboot <- 1000
alpha <- .1
lo <- nboot*(alpha/2)
hi <- nboot - lo
lo <- lo + 1

nexp <- 50 # number of experiments
n <- 20 # sample size in each experiment
n.m <- 50 # mean of normal distribution 
n.sd <- 20 # sd of normal distribution
ci <- matrix(NA, nrow = nexp, ncol = 2)
ci.ttest <- matrix(NA, nrow = nexp, ncol = 2)
for(E in 1:nexp){
  # sample data + bootstrap + compute mean of each bootstrap sample
  samp <- rnorm(n, n.m, n.sd)
  boot.samp <- apply(matrix(sample(samp, n*nboot, replace = TRUE), nrow = nboot), 1, mean)
  sort.boot.samp <- sort(boot.samp) # sort bootstrap estimates
  ci[E,1] <- sort.boot.samp[lo] 
  ci[E,2] <- sort.boot.samp[hi]
  ci.ttest[E,] <- t.test(samp, conf.level = 1-alpha)$conf.int
}
```

### Illustrate results: bootstrap
```{r, fig.height=3, fig.width=2}
df <- tibble(x = as.vector(ci),
             y = rep(1:nexp,2),
             gr = factor(rep(1:nexp,2))
            )
ggplot(df, aes(x = x, y = y)) + theme_linedraw() +
  geom_point() +
  geom_line(aes(group = gr)) +
  geom_vline(xintercept = n.m, colour = "red") +
  labs(x = "Values", y = "Experiments") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))

# https://stackoverflow.com/questions/12253239/vertical-lines-between-points-with-ggplot2
# ggsave(filename = './50expt_bootci.pdf')
```

### Illustrate results: t-test
```{r, fig.height=3, fig.width=2}
df <- tibble(x = as.vector(ci.ttest),
             y = rep(1:nexp,2),
             gr = factor(rep(1:nexp,2))
            )
ggplot(df, aes(x = x, y = y)) + theme_linedraw() +
  geom_point() +
  geom_line(aes(group = gr)) +
  geom_vline(xintercept = n.m, colour = "red") +
  labs(x = "Values", y = "Experiments") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
# ggsave(filename = './50expt_ttci.pdf')
```

