---
title: "Notebook 1: Introduction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Rand Wilcox's code
When I mention code from Rand Wilcox, it is available in a giant text file.
You access the functions by:
```{r, eval=FALSE}
source("./code/Rallfun-v35.txt")
```

See more detail and updated versions of the code [here](https://dornsife.usc.edu/labs/rwilcox/software/).

Some of the functions are also available in the `WRS2` package.

# Dice exercise

## Data capture

Enter the data in R or a text file: 5 rows (trials) and 10 columns (participants).

```{r}
# enter data in matrix with 5 rows, 10 columns:

# dice_res[,1] <- c(,,,,)
# dice_res[,2] <- c(,,,,)
# dice_res[,3] <- c(,,,,)
# dice_res[,4] <- c(,,,,)
# dice_res[,5] <- c(,,,,)
# dice_res[,6] <- c(,,,,)
# dice_res[,7] <- c(,,,,)
# dice_res[,8] <- c(,,,,)
# dice_res[,9] <- c(,,,,)
# dice_res[,10] <- c(,,,,)

# Save in a txt file:

# write.table(dice_res, file = "./data/dice_res_blue.txt", row.names=FALSE, col.names=FALSE)
# write.table(dice_res, file = "./data/dice_res_green.txt", row.names=FALSE, col.names=FALSE)
# write.table(dice_res, file = "./data/dice_res_red.txt", row.names=FALSE, col.names=FALSE)
```

## Load simulated data

```{r}
nt <- 5
np <- 10
res_red <- matrix(scan("./data/res1.txt"), nrow=nt, byrow=TRUE)
res_green <- matrix(scan("./data/res2.txt"), nrow=nt, byrow=TRUE)
res_blue <- matrix(scan("./data/res3.txt"), nrow=nt, byrow=TRUE)
```

## Load real data

```{r, eval = FALSE}
res_red <- matrix(scan("./data/dice_res_red.txt"), nrow=nt, byrow=TRUE)
res_green <- matrix(scan("./data/dice_res_green.txt"), nrow=nt, byrow=TRUE)
res_blue <- matrix(scan("./data/dice_res_blue.txt"), nrow=nt, byrow=TRUE)
```

## Illustration

Mean results
```{r}
m.red <- round(apply(res_red, 2, mean), digits=1)
m.green <- round(apply(res_green, 2, mean), digits=1)
m.blue <- round(apply(res_blue, 2, mean), digits=1)
```

Illustrate results
```{r}
set.seed(1)
df <- tibble(obs = c(m.red, m.green, m.blue),
             gp = factor(c(rep("Red",np),rep("Green",np),rep("Blue",np)))
             )

# preserve factor order
df$gp <- as.character(df$gp)
df$gp <- factor(df$gp, levels=unique(df$gp))

  ggplot(df, aes(gp, obs)) + theme_linedraw() +
  geom_point(aes(colour = gp), position = position_jitter(width = 0.1)) +
    scale_color_manual(values=c("red", "green4", "blue")) +
  theme(legend.position = "none",
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16)) +
  labs(x = "Groups", y = "Observations") +   
    coord_cartesian(ylim = c(0, 12)) +
    scale_y_continuous(breaks = seq(0, 12, 2)) +
  geom_errorbar(stat = "summary", fun.y = mean, fun.ymin = mean,
                fun.ymax = mean, colour = "grey", width = .55, size = 3) +
  geom_errorbar(stat = "summary", fun.y = mean, fun.ymin = mean,
                fun.ymax = mean, colour = "black", width = .5, size = 1)
```

## Inferential statistics?

What would you do next with the data?

