---
title: "Notebook 4: Sampling distributions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# dependencies
library(ggplot2)
library(tibble)
```

# Sampling distributions

## Sampling from a normal distribution

### Population
```{r}
m <- 50 # population mean
sd <- 10 # population standard deviation

ggplot(data.frame(x = c(0, 100)), aes(x)) + theme_linedraw() +
  stat_function(fun = dnorm, args = list(mean = m, sd = sd)) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("Normal distribution: mean = ",m,", sd = ",sd))
# ggsave(filename = "./figures/nb4_normdist.pdf")
```

### One sample
```{r}
set.seed(4)
n <- 20 # sample size
samp <- rnorm(n, m, sd) # generate random sample

ggplot(data.frame(x = c(0, 100)), aes(x)) + theme_linedraw() +
  stat_function(fun = dnorm, args = list(mean = m, sd = sd)) +
  geom_vline(xintercept = samp, colour = "grey") + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("Normal distribution: mean = ",m,", sd = ",sd))
# ggsave(filename = "./figures/nb4_normdist_1samp.pdf")
```

### One sample + mean

Same figure as above with the sample mean superimposed.

```{r}
set.seed(4)
n <- 20
samp <- rnorm(n, m, sd)
ggplot(data.frame(x = c(0, 100)), aes(x)) + theme_linedraw() +
  stat_function(fun = dnorm, args = list(mean = m, sd = sd)) +
  geom_vline(xintercept = samp, colour = "grey") +
  geom_vline(xintercept = mean(samp), size = 2) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste("Sample mean =",round(mean(samp),digits=1),", sd = ",round(sd(samp),digits=1)))
# ggsave(filename = "./figures/nb4_normdist_1samp_mean.pdf")
```

### Sampling distribution

Now we do 10,000 experiments, each time sampling from the same normal distribution and computing the sample mean. The graph below illustrates 10,000 sample means.

```{r}
set.seed(111) 
n <- 20 # sample size
nsim <- 10000 # number of simulations / experiments
# generate sampling distribution in one go:
samp.dist <- apply(matrix(rnorm(n*nsim, m, sd),nrow = nsim), 1, mean)

ggplot(data.frame(x=samp.dist), aes(x)) + theme_linedraw() +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white", bins = 50) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste(nsim,"sample means: mean =",round(mean(samp.dist),digits=1),", sd = ",round(sd(samp.dist),digits=2)))
# ggsave(filename = "./figures/nb4_normdist_10000samp_mean.pdf")
```

### Standard error of the mean

SEM is defined as: population sd / sqrt(n)

```{r}
sd / sqrt(n)
```
In practice, we do not have access to the population, so SEM is usually estimated using sample sd / sqrt(n).

SEM can also be defined as the sd of the sampling distribution of the sample mean.

```{r}
sd(samp.dist)
```

This value is very close to the previous one. If we increase the size of sampling distribution, by computing 100,000 or 1,000,000 or more sample means, the standard deviation of the sampling distribution will converge towards 2.236 precisely.

For more on the SEM, see this excellent [video](https://www.khanacademy.org/math/ap-statistics/sampling-distribution-ap/sampling-distribution-mean/v/standard-error-of-the-mean).

## Example using the lognormal

[Blog post](https://msalganik.wordpress.com/2017/01/21/making-sense-of-the-rlnorm-function-in-r/) on how to use `rlnorm()`.

### Illustrate lognormal
```{r}
# population
m <- 5
s <- 5
location <- log(m^2 / sqrt(s^2 + m^2))
shape <- sqrt(log(1 + (s^2 / m^2)))
x <- seq(0, 20, .05)
ggplot(tibble::enframe(x), aes(value)) +
  theme_linedraw() +
  stat_function(fun = dlnorm, args = list(meanlog = location, sdlog = shape), size = 1) +
  labs(y = "Density") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("Log normal distribution: mean = ",m,", sd = ",s))
# ggsave(filename = "./figures/nb4_lnormdist.pdf")
```

### One sample from lognormal distribution 

```{r}
set.seed(777)
n <- 20
samp <- rlnorm(n, meanlog = location, sdlog = shape)
x <- seq(0, 20, .05)
ggplot(as_tibble(x), aes(value)) +
  theme_linedraw() +
  stat_function(fun = dlnorm, args = list(meanlog = location, sdlog = shape), size = 1) +
  geom_vline(xintercept = samp, colour = "grey") + 
  geom_vline(xintercept = mean(samp), size = 1.5) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("Log normal distribution: mean = ",m,", sd = ",sd))
# ggsave(filename = "./figures/nb4_lnormdist_1samp.pdf")
```

### Many samples from lognormal distribution 

```{r}
set.seed(777)
m <- 5
s <- 5
location <- log(m^2 / sqrt(s^2 + m^2))
shape <- sqrt(log(1 + (s^2 / m^2)))

# population mean?
lnorm.pop.m <- mean(rlnorm(1000000, meanlog = location, sdlog = shape))

n <- 20
nsamp <- 10000
# nsamp experiments in one go + compute mean for each one:
dist.samp <- apply(matrix(rlnorm(n*nsamp, meanlog = location, sdlog = shape), nrow = nsamp), 1, mean)

v <- as_tibble(dist.samp) 
ggplot(v, aes(x = value)) + theme_linedraw() +
        geom_histogram(aes(y = ..density..), colour = "black", fill = "white", bins = 50) +
  # geom_vline(xintercept = mean(dist.samp), size = 1.5) + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste(nsim,"sample means: mean =",round(mean(dist.samp),digits=1),", sd = ",round(sd(dist.samp),digits=2)))
# ggsave(filename = "./figures/nb4_lnormdist_10000samp_mean.pdf")
```

### SEM

Based on (otherwise unobservable) population:
```{r}
s / sqrt(n)
```

Based on (otherwise unachievable) 10,000 experiments:
```{r}
sd(dist.samp)
```

Based on one sample:
```{r}
sd(samp) / sqrt(n)
```

# Bootstrap sampling distribution

## Example 1: samples from a standard normal distribution

### Sampling distribution
```{r}
set.seed(777)
n <- 100 # sample size
nsamp <- 10000 # number of samples (experiments)
# sampling distribution of the mean
dist.samp <- apply(matrix(rnorm(n*nsamp), nrow = nsamp), 1, mean)
v <- enframe(dist.samp, name = NULL) 
ggplot(v, aes(x = value)) +
        geom_histogram(aes(y = ..density..), 
                       bins = 40,  colour = "black", fill = "white") +
        geom_line(aes(y = ..density.., colour = 'Sampling'), stat = 'density', size = 1) +     
        scale_colour_manual(name = "Distributions", values = "orange1") +
  theme(legend.position = c(0.1, 0.85),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)
       )
# ggsave(filename = './figures/nb4_norm_samp_dist.pdf')
```

### Sampling distribution + bootstrap distribution [1]

The bootstrap distribution can approximate the sampling distribution very well for n = 100 and nboot = 1000. Comment out set.seed, so that different random numbers are generated each time you run the chunk.
What happens for different random samples?

What happens for lower values of n and nboot? 
What do you conclude?

```{r}
set.seed(777)
n <- 100 # sample size 
samp <- rnorm(n) # create sample from normal population
nboot <- 1000 # number of bootstrap samples
nsamp <- 10000 # number of samples (experiments)
# bootstrap distribution
boot.samp <- apply(matrix(sample(samp, n*nboot, replace = TRUE), nrow = nboot), 1, mean)
# sampling distribution of the mean
dist.samp <- apply(matrix(rnorm(n*nsamp), nrow = nsamp), 1, mean)
v <- enframe(dist.samp, name = NULL) 
ggplot(v, aes(x = value)) +
        geom_histogram(aes(y = ..density..), 
                       bins = 40,  colour = "black", fill = "white") +
        geom_line(aes(y = ..density.., colour = 'Sampling'), stat = 'density', size = 1) +     
        # stat_function(fun = dnorm, aes(colour = 'Normal'),
        #                  args = list(mean = mean(boot.samp), sd = sd(boot.samp)), size = 1) +
        geom_line(data = as_tibble(boot.samp),  
                  aes(x = value, y = ..density.., colour = 'Bootstrap'), 
                  stat = 'density', size = 1) +
        scale_colour_manual(name = "Distributions", values = c("green4", "orange1")) +
  theme(legend.position = c(0.1, 0.85),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        legend.text = element_text(size = 12),
      legend.title = element_text(size = 14)
       )
# ggsave(filename = './figures/nb4_norm_samp_dist_boot.pdf')
```

### Sampling distribution + bootstrap distribution [2]

```{r}
set.seed(666)
n <- 100 # sample size 
samp <- rnorm(n) # create sample from normal population
nboot <- 1000 # number of bootstrap samples
nsamp <- 10000 # number of samples (experiments)
# bootstrap distribution
boot.samp <- apply(matrix(sample(samp, n*nboot, replace = TRUE), nrow = nboot), 1, mean)
# sampling distribution of the mean
dist.samp <- apply(matrix(rnorm(n*nsamp), nrow = nsamp), 1, mean)
v <- enframe(dist.samp, name = NULL) 
ggplot(v, aes(x = value)) +
        geom_histogram(aes(y = ..density..), 
                       bins = 40,  colour = "black", fill = "white") +
        geom_line(aes(y = ..density.., colour = 'Sampling'), stat = 'density', size = 1) +     
        # stat_function(fun = dnorm, aes(colour = 'Normal'),
        #                  args = list(mean = mean(boot.samp), sd = sd(boot.samp)), size = 1) +
        geom_line(data = as_tibble(boot.samp),  
                  aes(x = value, y = ..density.., colour = 'Bootstrap'), 
                  stat = 'density', size = 1) +
        scale_colour_manual(name = "Distributions", values = c("green4", "orange1")) +
  theme(legend.position = c(0.1, 0.85),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        legend.text = element_text(size = 12),
      legend.title = element_text(size = 14)
       )
# ggsave(filename = './figures/nb4_norm_samp_dist_boot2.pdf')
```

## Example 2: sample from gamma distribution

### Check shape of gamme distribution
```{r}
x <- seq(0, 15, .05)
ggplot(tibble::enframe(x), aes(value)) +
  theme_linedraw() +
  stat_function(fun = dgamma, args = list(shape = 2, scale = 2), size = 1) +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14))
# ggsave(filename = './figures/nb4_gamma_dist.pdf')
```

### Sampling distribution + bootstrap distribution [1]

```{r}
set.seed(777) 
n <- 100
gamma_shape <- 2
gamma_scale <- 2
# sample n trials from gamma distribution
samp <- rgamma(n, shape = gamma_shape, scale = gamma_scale)
nboot <- 1000
nsamp <- 10000
# bootstrap distribution
boot.samp <- apply(matrix(sample(samp, n*nboot, replace = TRUE), nrow = nboot), 1, mean)
# sampling distribution of the mean of n trials from gamma distribution
dist.samp <- apply(matrix(rgamma(n*nsamp, shape = gamma_shape, scale = gamma_scale), nrow = nsamp), 1, mean)
v <- enframe(dist.samp, name = NULL) 
ggplot(v, aes(x = value)) +
        geom_histogram(aes(y = ..density..), 
                       bins = 40,  colour = "black", fill = "white") +
        geom_line(aes(y = ..density.., colour = 'Sampling'), stat = 'density', size = 1) +     
        # stat_function(fun = dnorm, aes(colour = 'Normal'),
        #                  args = list(mean = mean(boot.samp), sd = sd(boot.samp)), size = 1) +
        geom_line(data = as_tibble(boot.samp),  
                  aes(x = value, y = ..density.., colour = 'Bootstrap'), 
                  stat = 'density', size = 1) +
        scale_colour_manual(name = "Distributions", values = c("green4", "orange1")) +
  theme(legend.position = c(0.1, 0.85),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        legend.text = element_text(size = 12),
      legend.title = element_text(size = 14)
       )
# ggsave(filename = './figures/nb4_gamma_samp_dist_boot.pdf')
```

### Sampling distribution + bootstrap distribution [2]

```{r}
set.seed(666) 
n <- 100
gamma_shape <- 2
gamma_scale <- 2
# sample n trials from gamma distribution
samp <- rgamma(n, shape = gamma_shape, scale = gamma_scale)
nboot <- 1000
nsamp <- 10000
# bootstrap distribution
boot.samp <- apply(matrix(sample(samp, n*nboot, replace = TRUE), nrow = nboot), 1, mean)
# sampling distribution of the mean of n trials from gamma distribution
dist.samp <- apply(matrix(rgamma(n*nsamp, shape = gamma_shape, scale = gamma_scale), nrow = nsamp), 1, mean)
v <- enframe(dist.samp, name = NULL) 
ggplot(v, aes(x = value)) +
        geom_histogram(aes(y = ..density..), 
                       bins = 40,  colour = "black", fill = "white") +
        geom_line(aes(y = ..density.., colour = 'Sampling'), stat = 'density', size = 1) +     
        # stat_function(fun = dnorm, aes(colour = 'Normal'),
        #                  args = list(mean = mean(boot.samp), sd = sd(boot.samp)), size = 1) +
        geom_line(data = as_tibble(boot.samp),  
                  aes(x = value, y = ..density.., colour = 'Bootstrap'), 
                  stat = 'density', size = 1) +
        scale_colour_manual(name = "Distributions", values = c("green4", "orange1")) +
  theme(legend.position = c(0.1, 0.85),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        legend.text = element_text(size = 12),
      legend.title = element_text(size = 14)
       )
# ggsave(filename = './figures/nb4_gamma_samp_dist_boot2.pdf')
```