---
title: "Notebook 4: Sampling distributions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# dependencies
library(ggplot2)
library(tibble)
```

```{r}
# population
m <- 50
sd <- 10
ggplot(data.frame(x = c(0, 100)), aes(x)) + theme_linedraw() +
  stat_function(fun = dnorm, args = list(mean = m, sd = sd)) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("Normal distribution: mean = ",m,", sd = ",sd))
# ggsave(filename = "b3_normdist.pdf")
```

```{r}
# one sample
set.seed(4)
n <- 20
samp <- rnorm(n, m, sd)
ggplot(data.frame(x = c(0, 100)), aes(x)) + theme_linedraw() +
  stat_function(fun = dnorm, args = list(mean = m, sd = sd)) +
  geom_vline(xintercept = samp, colour = "grey") + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("Normal distribution: mean = ",m,", sd = ",sd))
# ggsave(filename = "b3_normdist_1samp.pdf")
```

```{r}
# one sample + mean
set.seed(4)
n <- 20
samp <- rnorm(n, m, sd)
ggplot(data.frame(x = c(0, 100)), aes(x)) + theme_linedraw() +
  stat_function(fun = dnorm, args = list(mean = m, sd = sd)) +
  geom_vline(xintercept = samp, colour = "grey") +
  geom_vline(xintercept = mean(samp), size = 2) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste("Sample mean =",round(mean(samp),digits=1),", sd = ",round(sd(samp),digits=1)))
# ggsave(filename = "b3_normdist_1samp_mean.pdf")
```

```{r}
# sampling distribution
set.seed(111) 
n <- 20
nsim <- 10000
samp.dist <- apply(matrix(rnorm(n*nsim, m, sd),nrow = nsim), 1, mean)

ggplot(data.frame(x=samp.dist), aes(x)) + theme_linedraw() +
  geom_histogram(aes(y = ..density..), colour = "black", fill = "white", bins = 50) +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste(nsim,"sample means: mean =",round(mean(samp.dist),digits=1),", sd = ",round(sd(samp.dist),digits=2)))
# ggsave(filename = "b3_normdist_10000samp_mean.pdf")
```

# Standard error of the mean

SEM = population sd / sqrt(n)

Usually estimated using sample sd / sqrt(n)

```{r}
sd / sqrt(n)
```
SEM = sd of sampling distribution of the sample mean

https://www.khanacademy.org/math/ap-statistics/sampling-distribution-ap/sampling-distribution-mean/v/standard-error-of-the-mean

```{r}
sd(samp.dist)
```

# t-test confidence interval

## Get one sample and apply t.test() function
```{r}
# one sample
set.seed(4)
n <- 20
samp <- rnorm(n, m, sd)
t.test(samp)
```

## Compute t value
```{r}
# Formula
samp.m <- mean(samp) # sample mean
samp.v <- var(samp) # sample variance
sem <- sqrt(samp.v/n) # standard error of the mean
samp.t <- samp.m / sem
samp.t
```

## t distribution and quantiles
```{r}
ggplot(data.frame(x = c(-5, 5)), aes(x)) + theme_linedraw() +
  stat_function(fun = dt, args = list(df = df)) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("t distribution: df = ",df,", ",(1-alpha/2)*100,"th quantile = ",round(qt(1-alpha/2, df), digits=2))) +
  geom_vline(xintercept = c(-qt(1-alpha/2, df), qt(1-alpha/2, df)), colour = "green", size = 1)
```

## Compute confidence interval
```{r}
alpha <- 0.05
df <- n-1 # degrees of freedom
ci <- vector("numeric",2)
ci[1] <- samp.m - qt(1-alpha/2, df)  * sem
ci[2] <- samp.m + qt(1-alpha/2, df)  * sem
ci
```

# Example using the lognormal

[Blog post](https://msalganik.wordpress.com/2017/01/21/making-sense-of-the-rlnorm-function-in-r/) on how to use `rlnorm()`.

## Illustrate lognormal
```{r}
# population
m <- 5
s <- 5
location <- log(m^2 / sqrt(s^2 + m^2))
shape <- sqrt(log(1 + (s^2 / m^2)))
x <- seq(0, 20, .05)
ggplot(as_tibble(x), aes(value)) +
  theme_linedraw() +
  stat_function(fun = dlnorm, args = list(meanlog = location, sdlog = shape), size = 1) +
  labs(y = "Density") +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("Log normal distribution: mean = ",m,", sd = ",sd))
# ggsave(filename = "b3_lnormdist.pdf")
```

## One sample from lognormal distribution 
```{r}
set.seed(777)
n <- 20
samp <- rlnorm(n, meanlog = location, sdlog = shape)
x <- seq(0, 20, .05)
ggplot(as_tibble(x), aes(value)) +
  theme_linedraw() +
  stat_function(fun = dlnorm, args = list(meanlog = location, sdlog = shape), size = 1) +
  geom_vline(xintercept = samp, colour = "grey") + 
  geom_vline(xintercept = mean(samp), size = 1.5) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("Log normal distribution: mean = ",m,", sd = ",sd))
# ggsave(filename = "b3_lnormdist_1samp.pdf")
```

## Many samples from lognormal distribution 
```{r}
set.seed(777)
m <- 5
s <- 5
location <- log(m^2 / sqrt(s^2 + m^2))
shape <- sqrt(log(1 + (s^2 / m^2)))

# population mean?
lnorm.pop.m <- mean(rlnorm(1000000, meanlog = location, sdlog = shape))

n <- 20
nsamp <- 10000
# nsamp experiments in one go + compute mean for each one:
dist.samp <- apply(matrix(rlnorm(n*nsamp, meanlog = location, sdlog = shape), nrow = nsamp), 1, mean)

v <- as_tibble(dist.samp) 
ggplot(v, aes(x = value)) + theme_linedraw() +
        geom_histogram(aes(y = ..density..), colour = "black", fill = "white", bins = 50) +
  # geom_vline(xintercept = mean(dist.samp), size = 1.5) + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste(nsim,"sample means: mean =",round(mean(dist.samp),digits=1),", sd = ",round(sd(dist.samp),digits=2)))
# ggsave(filename = "b3_lnormdist_10000samp_mean.pdf")
```

