---
title: "Notebook 5: Bootstrap-t"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tibble)
```

# t-test confidence interval: example 1

## Get one sample and apply t.test() function
```{r}
# one sample
set.seed(4)
n <- 20
m <- 0
sd <- 1
samp <- rnorm(n, m, sd)
t.test(samp)
```

## Compute t value
```{r}
# Formula
samp.m <- mean(samp) # sample mean
samp.v <- var(samp) # sample variance
sem <- sqrt(samp.v/n) # standard error of the mean
samp.t <- samp.m / sem
samp.t
```

## Compute P value
```{r}
df <- n-1
samp.pval <- 2*(1-pt(abs(samp.t),df))
samp.pval
```

### Illustrate P value
```{r}
alpha <- 0.05

# df <- as_tibble(with(density(boot.m),data.frame(x,y)))
# 
# df.pv1 <- tibble(x = df$x[df$x<nv],
#                 y = df$y[df$x<nv])
# 
# df.pv2 <- tibble(x = df$x[df$x>nv],
#                 y = df$y[df$x>nv])

df.pv2 <- tibble(x = seq(samp.t,5,0.1),
                 y = dt(seq(samp.t,5,0.1), df))

ggplot(data.frame(x = c(-5, 5)), aes(x)) + theme_linedraw() +
  stat_function(fun = dt, args = list(df = df)) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("t distribution: df = ",df,", t value = ",round(samp.t, digits=2))) +
  geom_area(data = df.pv2,
            aes(x = x, y = y),
            fill = "red", alpha = .2) +
  geom_vline(xintercept = samp.t, colour = "red", size = 1)
  # geom_vline(xintercept = samp.t, colour = "purple", size = 1)
```

## t distribution and quantiles

Here we illustrate the alpha/2 and 1-alpha/2 quantiles of the t distribution with 19 degrees of freedom, under the null (given that there is no effect).

```{r}
alpha <- 0.05

ggplot(data.frame(x = c(-5, 5)), aes(x)) + theme_linedraw() +
  stat_function(fun = dt, args = list(df = df)) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("t distribution: df = ",df,", ",(1-alpha/2)*100,"th quantile = ",round(qt(1-alpha/2, df), digits=2))) +
  geom_vline(xintercept = c(-qt(1-alpha/2, df), qt(1-alpha/2, df)), colour = "green", size = 1)
  # geom_vline(xintercept = samp.t, colour = "purple", size = 1)
```

## Compute confidence interval
```{r}
alpha <- 0.05
df <- n-1 # degrees of freedom
ci <- vector("numeric",2)
ci[1] <- samp.m - qt(1-alpha/2, df)  * sem
ci[2] <- samp.m + qt(1-alpha/2, df)  * sem
ci
```

# t-test confidence interval: example 2

## Get one sample and apply t.test() function
```{r}
# one sample
set.seed(21)
n <- 20
m <- 0
sd <- 1
samp <- rnorm(n, m, sd)
t.test(samp)
```

## Compute t value
```{r}
# Formula
samp.m <- mean(samp) # sample mean
samp.v <- var(samp) # sample variance
sem <- sqrt(samp.v/n) # standard error of the mean
samp.t <- samp.m / sem
samp.t
```

## Compute P value
```{r}
df <- n-1 # degrees of freedom
samp.pval <- 2*(1-pt(abs(samp.t),df))
samp.pval
```

## t distribution and quantiles
```{r}
alpha <- 0.05 # arbitrary significance threshold

ggplot(data.frame(x = c(-5, 5)), aes(x)) + theme_linedraw() +
  stat_function(fun = dt, args = list(df = df)) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("t distribution: df = ",df,", ",(1-alpha/2)*100,"th quantile = ",round(qt(1-alpha/2, df), digits=2))) +
  geom_vline(xintercept = c(-qt(1-alpha/2, df), qt(1-alpha/2, df)), colour = "green", size = 1) +
  geom_vline(xintercept = samp.t, colour = "purple", size = 1)
```

## Compute confidence interval
```{r}
ci <- vector("numeric",2)
ci[1] <- samp.m - qt(1-alpha/2, df)  * sem
ci[2] <- samp.m + qt(1-alpha/2, df)  * sem
ci
```

# t-test confidence interval: example 3

## Get one sample and apply t.test() function
```{r}
# one sample
set.seed(7)
n <- 20
m <- 0.5
sd <- 1
samp <- rnorm(n, m, sd)
t.test(samp)
```

## Compute t value
```{r}
# Formula
samp.m <- mean(samp) # sample mean
samp.v <- var(samp) # sample variance
sem <- sqrt(samp.v/n) # standard error of the mean
samp.t <- samp.m / sem
samp.t
```

## Compute P value
```{r}
df <- n-1 # degrees of freedom
samp.pval <- 2*(1-pt(abs(samp.t),df))
samp.pval
```

## t distribution and quantiles
```{r}
alpha <- 0.05 # arbitrary significance threshold

ggplot(data.frame(x = c(-5, 5)), aes(x)) + theme_linedraw() +
  stat_function(fun = dt, args = list(df = df)) + 
labs(y = "Density") +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size=20)) +
  ggtitle(paste0("t distribution: df = ",df,", ",(1-alpha/2)*100,"th quantile = ",round(qt(1-alpha/2, df), digits=2))) +
  geom_vline(xintercept = c(-qt(1-alpha/2, df), qt(1-alpha/2, df)), colour = "green", size = 1) +
  geom_vline(xintercept = samp.t, colour = "purple", size = 1)
```

## Compute confidence interval
```{r}
ci <- vector("numeric",2)
ci[1] <- samp.m - qt(1-alpha/2, df)  * sem
ci[2] <- samp.m + qt(1-alpha/2, df)  * sem
ci
```

# Bootstrap-t technique

See detailed explanations and R code in this [tutorial](https://psyarxiv.com/h8ft7/).
