---
title: "Notebook 7: Perspectives on group comparisons"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tibble)
source("./code/theme_gar.txt")
library(ggbeeswarm)
source("./code/Rallfun-v35-simplified.txt")
source("./code/summarySE.txt") # Define summary function
# Code from: 
# http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/
```

# Example: within x between design
The example is from a 2016 editorial published in the [European Journal of Neuroscience](https://onlinelibrary.wiley.com/doi/full/10.1111/ejn.13400).

## Load data
```{r, load_ejn_data}
data <- read.csv("./data/ejn_data.csv",header=T)
data$participant <- factor(data$participant)
```

## Bar graph
```{r, bar_graph}
datas <- summarySE(data, measurevar="value", groupvars=c("group","condition"))

bargraph <- ggplot(datas, aes(x=group, y=value, fill=condition)) + 
  theme_bw() +
  geom_bar(position=position_dodge(), stat="identity", colour="black") +
  geom_errorbar(aes(ymin=value-se, ymax=value+se),
                width=.2,                    # Width of the error bars
                position=position_dodge(.9)) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
theme(axis.text.x = element_text(colour="grey20",size=16),
      axis.text.y = element_text(colour="grey20",size=16),  
      axis.title.x = element_blank(),
      axis.title.y = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(colour="grey20",size=16),
      legend.key = element_rect(colour = "black",size=.1),
      legend.position = c(.15, .90),
      # plot.margin = unit(c(150,100,5.5,5.5), "pt"),
      plot.margin = unit(c(5.5,5.5,5.5,5.5), "pt"),
      plot.title = element_text(colour="grey20",size=20)) +
      labs(title="Mean +/- SEM") +
      scale_y_continuous(limits=c(0, 15),breaks=seq(0,15,5))
bargraph
```

## 1D scatterplots = stripcharts of differences 
```{r}
value <- data$value[data$condition=="condition2"]-data$value[data$condition=="condition1"]
participant <- data$participant[data$condition=="condition2"]
group <- data$group[data$condition=="condition2"]
diff <- data.frame(participant,group,value)

set.seed(8)

diffstrip <- ggplot(diff, aes(x=group, y=value, fill=group, colour=group, shape=group)) + theme_bw() +
  geom_abline(intercept = 0, slope = 0) +
  geom_jitter(position=position_jitter(0.1), size=4, stroke=1, alpha=0.6) +
  scale_shape_manual(values=c(22,21)) +
  scale_fill_manual(values = c("grey70", "grey95")) +
  scale_colour_manual(values = c("grey5","grey5")) +
  theme(legend.position="none",
        axis.text.x = element_text(colour="grey20",size=16),
        axis.text.y = element_text(colour="grey20",size=16),  
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(colour="grey20",size=20),
        plot.margin = unit(c(5.5,5.5,5.5,5.5), "pt")) +
        # plot.margin = unit(c(150,200,5.5,5.5), "pt")
  labs(title="Differences: condition1 - condition2") +
  scale_y_continuous(limits=c(-2.5, 3.5),breaks=seq(-2,3,1)) 
diffstrip
```

## Stripchart + linked observations 
```{r}
p <- ggplot(data, aes(x=condition, y=value, fill=condition, group=participant,shape=group)) + theme_bw() +
  geom_line(size=1, alpha=1) +
  geom_point(colour = "black", size = 4, stroke = 1) +
  scale_shape_manual(values=c(22,21)) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  theme(axis.text.x = element_text(colour="grey20",size=16),
        axis.text.y = element_text(colour="grey20",size=16),  
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title = element_text(colour="grey20",size=20),
        legend.position="none",
        # plot.margin = unit(c(150,100,5.5,5.5), "pt")
        plot.margin = unit(c(5.5,5.5,5.5,5.5), "pt")) +
        labs(title="Paired observations") +
  scale_y_continuous(limits=c(0, 20),breaks=seq(0,20,5))
# p
linkedstrip <- p + facet_grid(. ~ group) +
  theme(strip.text.x = element_text(size = 20, colour = "white"),
        strip.background = element_rect(colour="darkgrey", fill="darkgrey"))
linkedstrip
```

## Scatterplot of paired observations 
```{r}
condition1 <- data$value[data$condition=="condition1"]
condition2 <- data$value[data$condition=="condition2"]
participant <- data$participant[data$condition=="condition2"]
group <- data$group[data$condition=="condition2"]
rdata <- data.frame(participant,group,condition1,condition2)

scatterdiff <- ggplot(rdata, aes(x=condition1,y=condition2,group=group,fill=group,colour=group,shape=group)) + 
  geom_abline(intercept = 0) +
  geom_point(size=4,stroke=1, alpha = 0.6) +
  theme_bw() +
  scale_shape_manual(values=c(22,21)) +
  scale_fill_manual(values = c("grey70", "grey95")) +
  scale_colour_manual(values = c("grey5","grey5")) +
  theme(axis.text.x = element_text(colour="grey20",size=16),
        axis.text.y = element_text(colour="grey20",size=16),  
        axis.title.x = element_text(colour="grey20",size=18),
        axis.title.y = element_text(colour="grey20",size=18),
        legend.title = element_blank(),
        legend.position = c(.15, .90),
        plot.margin = unit(c(5.5,5.5,5.5,5.5), "pt"),
          legend.text = element_text(colour="grey20",size=16),
        plot.title = element_text(colour="grey20",size=20)) +
  labs(title="Paired observations") +
  scale_x_continuous(limits=c(6, 16),breaks=seq(6,16,2)) +
  scale_y_continuous(limits=c(6, 19),breaks=seq(6,19,2))
scatterdiff
```

## Examples of standard group statistics

Stats results as reported in the EJN editorial.

### 2 x 2 ANOVA

```{r}
aov_gp_cond <- aov(value ~ group*condition + Error(participant/condition), data=data)
summary(aov_gp_cond)
```

### Group 1 t-test
```{r}
x <- data[data$group=="group1" & data$condition=="condition1" ,"value"]
y <- data[data$group=="group1" & data$condition=="condition2" ,"value"]
t.test(x,y,paired=T)
```

### Group 2 t-test
```{r}
x <- data[data$group=="group2" & data$condition=="condition1" ,"value"]
y <- data[data$group=="group2" & data$condition=="condition2" ,"value"]
t.test(x,y,paired=T)
```

## Make final figure combining 4 panels
```{r, eval=FALSE}
require(cowplot)
set.seed(8)
cowplot::plot_grid(bargraph, diffstrip, linkedstrip, scatterdiff, 
                   labels=c("A", "B", "C", "D"), 
                   ncol = 2, 
                   nrow = 2, 
                   rel_heights = c(1, 1),
                   label_size = 18,
                   hjust = -1,
                   scale=.95)

ggsave(filename='./figures/ejn_summary_figure.pdf', units = "cm", width = 30, height = 30)
```

# Two dependent groups

Here are two vectors of ages of husbands and wives. The two vectors are matching, such that husbands[1] is married to wives[1].

```{r}
husbands <- c(25,25,51,25,38,30,60,54,31,54,23,34,25,23,19,71,26,31,26,62,29,31,29,35)

wives <- c(22,32,50,25,33,27,45,47,30,44,23,39,24,22,16,73,27,36,24,60,26,23,28,36)

n.couple <- length(wives)
```

## Illustrate results

### Make data frame
```{r}
df <- tibble(couple = rep(1:n.couple, 2),
             age = c(husbands, wives),
             group = rep(c("husbands", "wives"), each = n.couple))
```

### Independent stripcharts (1D scatterplots)
```{r}
set.seed(44) # reproducible jitter

ggplot(df, aes(x = group,
               y = age,
               fill = group)) +
geom_jitter(alpha = 0.7,
            shape = 21,
            size = 4,
            width = .15) +
  scale_fill_manual(values=c("#E69F00", "#56B4E9")) +
    theme_bw() +
    theme(legend.position = "none") +
    theme(axis.title.y = element_text(size=16,face="bold"),
        axis.title.x = element_blank(),
        axis.text = element_text(size=14)) +
    ylab("Ages in years") +
  scale_y_continuous(limits = c(0, 100))
```

### Stripcharts with medians

```{r}
set.seed(44) # reproducible jitter

# compute medians
md.h <- median(husbands)
md.w <- median(wives)

ggplot(df, aes(x = group,
               y = age,
               fill = group)) +
geom_jitter(alpha = 0.7,
            shape = 21,
            size = 4,
            width = .15) +
scale_fill_manual(values=c("#E69F00", "#56B4E9")) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(size=16,face="bold"),
        axis.title.x = element_blank(),
        axis.text = element_text(size=14)) +
        ylab("Ages in years") +
scale_y_continuous(limits = c(0, 100)) +
# add median for husbands

geom_segment(aes(x = 0.75, y = md.h, xend = 1.25, yend = md.h), size = 1, colour = "black") +
# add median for wives
geom_segment(aes(x = 1.75, y = md.w, xend = 2.25, yend = md.w), size = 1, colour = "black")
```

### Linked stripcharts

Coloured lines link paired observations.
```{r}
set.seed(44) # reproducible jitter

pd <- position_dodge(width = 0.1)

ggplot(df, aes(x=group, y=age, group=couple)) + theme_bw() +
  geom_line(aes(colour=couple),size=1, alpha=.9,
            position=pd) +
  scale_colour_viridis_c(option = "A") +
  theme(axis.text = element_text(size=14),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=16,face="bold"),
        legend.position="none") +
  # ylab("Scores (a.u.)") +
  # scale_x_discrete(labels = c("condition1" = "Condition 1","condition2" = "Condition 2")) +
  scale_y_continuous(limits=c(0, 100)) +
  ylab("Ages in years")
```

### 1D scatterplot = stripchart of differences

```{r}
# fig.width=0.75, fig.height=1.5
set.seed(44) # reproducible jitter

df.diff <- tibble(couple = 1:n.couple,
                  difference = husbands - wives,
                  gr = rep("gr", n.couple))

ggplot(df.diff, aes(x=gr, y=difference)) +
  geom_abline(intercept = 0, slope = 0, linetype = 2) + # zero reference line
  geom_jitter(alpha = 0.7,
              shape = 21,
              size = 2,
              width = .15,
              colour = "grey10",
              fill = "grey90") +
  theme_bw() +
  theme(legend.position="none",
        axis.ticks.x = element_line(colour="white"),
        axis.text.x = element_text(size=14,colour="white"),
        axis.text.y = element_text(size=14),
        axis.title.x = element_text(size=16,face="bold",colour="white"),
        axis.title.y = element_text(size=16,face="bold"),
        plot.title = element_text(colour="black",size=20),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  ylab("Differences") +
  scale_y_continuous(limits=c(-10, 20), breaks=seq(-10, 20, 5))
```

### Scatterplot of paired observations
```{r}
# make new data frame for paired data
df.p <- tibble(husbands = husbands,
               wives = wives)

ggplot(df.p, aes(x = husbands, y = wives)) + theme_bw() +
  geom_point(alpha = 0.7,
              shape = 21,
              size = 4,
              colour = "grey10",
              fill = "grey90") +
  geom_abline(intercept = 0,
              slope = 1,
              linetype = "longdash") + # slope 1 reference line
  theme(axis.title = element_text(size=16,face="bold"),
        axis.text = element_text(size=14)) +
  xlab("Wives") +
  ylab("Husbands") +
  theme(plot.title = element_blank())
```

## Compare marginals

Treated as two independent groups, are men older than women?

```{r}
# percentile bootstrap confidence interval of the difference between
# medians, treating the groups as independent
pb2gen(husbands, wives, alpha=.05, nboot=2000, est=median)
```


## Quantify within couple differences

Can we conclude that, within a couple, the wife tends to be older than the husband?

```{r}
# Compute confidence interval of the median of the paired differences
onesampb(husbands - wives, alpha=.05, nboot=2000, est=median)
```

### Summary statistics
```{r}
summary(husbands - wives)
```

## $P(diff > 0)$
```{r}
cdiff <- husbands - wives
mean(cdiff > 0)
```
## $P(diff > 0)$ + confidence interval

### Define function
Probability than x is larger than zero.
```{r}
pxlz<- function(x){
  mean(x > 0)
}
```

### Percentile bootstrap
```{r}
onesampb(cdiff, alpha=.05, nboot=2000, est=pxlz)
```

# Two independent groups

A study assessed the effects of ozone on weight gain in rats. The ozone group consisted of 22 70-day old rats kept in an ozone environment for 7 days. The control group consisted of 23 rats of the same age, kept in an ozone-free environment. Here are the weight gains, in grams, for each group:

```{r}
control <- c(41, 38.4, 24.4, 25.9, 21.9, 18.3, 13.1, 27.3, 28.5, -16.9, 26, 17.4, 21.8, 15.4, 27.4, 19.2, 22.4, 17.7, 26, 29.4, 21.4, 26.6, 22.7)

ozone <- c(10.1, 6.1, 20.4, 7.3, 14.3, 15.5, -9.9, 6.8, 28.2, 17.9, -9, -12.9, 14, 6.6, 12.1, 15.7, 39.9, -15.9, 54.6, -14.7, 44.1, -9)

n.control <- length(control)
n.ozone <- length(ozone)
```

## Illustrate results

### Make data frame
```{r}
df <- tibble(weight.gain = c(control, ozone),
             group = c(rep("control", n.control), rep("ozone", n.ozone)))
```

### Stripcharts with medians

```{r}
set.seed(44) # reproducible jitter

ggplot(df, aes(x = group,
               y = weight.gain,
               fill = group)) +
geom_jitter(alpha = 0.7,
            shape = 21,
            size = 4,
            width = .15) +
scale_fill_manual(values=c("#E69F00", "#56B4E9")) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.title.y = element_text(size=16,face="bold"),
        axis.title.x = element_blank(),
        axis.text = element_text(size=14)) +
        ylab("Weight gain in grams") +
coord_cartesian(ylim = c(-30, 60)) +
   stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                 geom = "crossbar", width = 0.3) +
   stat_summary(fun.y = mean, fun.ymin = mean, fun.ymax = mean,
                 geom = "crossbar", width = 0.3, colour = "purple",
                 fun.args = list(trim = 0.2))
```

These illustrations suggest differences in central tendency and spread between the two groups, as well as overlap, in part due to one or more outliers in the control group. The control group seems overall to gain more weights, with less inter-animal variability than the ozone group.

## Compare 20% trimmed means
```{r}
out <- pb2gen(control, ozone, alpha=.05, nboot=2000,
              est=mean, trim = 0.2)
```

Difference between marginal trimmed means = `r round(out$est.dif, digits=1)`, [`r round(out$ci[1], digits=1)`, `r round(out$ci[2], digits=1)`], *P*=`r round(out$p.value, digits=3)`.

## All pairwise differences

### Compute
```{r}
apd <- as.vector(outer(control, ozone, FUN="-"))
```

### All pairwise differences: implementation
```{r}
x <- 1:5
y <- 1:5
outer(x,y,FUN="-")
```

### Illustrate
Make kernel density plot of all pairwise differences
```{r}
ggplot(tibble::enframe(apd), aes(value)) + theme_bw() +
  geom_density(alpha = 1 , fill = "grey95", colour = "black") +
  geom_vline(xintercept = 0, linetype = "dashed") + # zero reference line
  geom_vline(xintercept = median(apd)) + # median difference
  theme(legend.position="none",
        axis.text = element_text(size=14),
        axis.title = element_text(size=16,face="bold")) +
  ylab("Density") +
  xlab("All pairwise differences")
```

### Median of all pairwise differences

For a measure of effect size preserving the original units, we could report the median of all pairwise differences and its percentile bootstrap confidence interval:
```{r}
out <- wmwpb(control,ozone,est=median,alpha=.05,nboot=2000)
```

Median of all pairwise differences = `r out$estimate` [`r out$ci[1]`, `r out$ci[2]`], *P*=`r out$p.value`.

This measure answer the questions: what is the typical difference between two observations selected at random from the two groups?

### $P(X>Y)$

Given two observations selected at random from these two groups, what is the probability that an observation from group 1 will be larger than one in group 2?

Using the `cid()` function:
```{r}
out <- cid(control, ozone)
out$summary.dvals[3]
```

`cid()` also returns the complementary probability $P(X<Y)$ and the robust ordinal measure of effect size known as Cliff's delta: $d = P(X>Y) - P(X<Y)$.

Make our own function to compute the probability that X is greater than Y:
```{r}
pxgy <- function(x,y){
  mean(outer(x,y,FUN="-") > 0)
}
```

Use it:
```{r}
pxgy(control, ozone)
```

Confidence interval:
```{r}
set.seed(21)

nboot <- 2000
b.res <- vector(mode = "numeric", length = nboot)
for(B in 1:nboot){
  b.res[B] <- pxgy(sample(control, n.control, replace = TRUE),
                   sample(ozone, n.ozone, replace = TRUE))
}
quantile(b.res, probs = c(0.025, 0.975))
```

