---
title: "Notebook 8: Shift function"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(tibble)
library(retimes) # to generate exGaussian data
# install.packages("devtools")
# devtools::install_github("GRousselet/rogme")
library(rogme) # *** NOT ON CRAN, SO FOLLOW THE TWO PREVIOUS STEPS TO INSTALL ROGME ***
```

Here we consider differences in reaction time distributions, modelled using [exGaussian distributions](https://www.ejwagenmakers.com/2009/MatzkeWagenmakers2009.pdf).

We look at 4 types of differences:  
- uniform shift;  
- difference in spread;  
- early difference;
- late difference.

# Uniform shift

## KDE
```{r}
nt <- 1000
# ex Gaussian parameters
mu <- 500
sigma <- 50
tau <- 200
ES <- 50

set.seed(21)

g1 <- rexgauss(nt, mu = mu, sigma = sigma, tau = tau) 
g2 <- rexgauss(nt, mu = mu, sigma = sigma, tau = tau) + ES
df <- mkt2(g1, g2, group_labels = c("Condition1", "Condition2"))

p <- ggplot(df, aes(x = obs)) + theme_classic() + 
  stat_density(aes(colour = gr), geom="line",position="identity", size=1) +
  scale_colour_viridis_d(end = 0.8) +
  coord_cartesian(xlim = c(0, 2500)) +
  theme(axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 16, colour="black"),
        axis.text.y = element_text(size = 16, colour="black"),
        axis.title.y = element_text(size = 18),
        legend.key.width = unit(1.5,"cm"),
        legend.position = c(0.55,0.75),
        legend.direction = "vertical",
        legend.text=element_text(size=16),
        legend.title=element_text(size=18),
        title = element_text(size=20)) +
  labs(x = "Reaction times", y = "Density", colour = "Conditions") +
  ggtitle("Uniform shift")
p
p.uni.dist <- p
```

## Shift function
```{r}
out <- shifthd_pbci(df, nboot = 200, adj_ci = FALSE)
p <- plot_sf(out, plot_theme = 1)[[1]] + 
     theme(axis.text = element_text(size = 16, colour="black"))
p
p.uni.sf <- p
```

# Difference in spread

## KDE
```{r}
nt <- 1000
# ex Gaussian parameters
mu <- 500
sigma <- 50
tau <- 200
ES <- 1.5

set.seed(777)
g1 <- rexgauss(nt, mu = mu, sigma = sigma, tau = tau)
md.g1 <- median(g1)
g1 <- (g1 - md.g1) * ES + md.g1 # spread around the median
g2 <- rexgauss(nt, mu = mu, sigma = sigma, tau = tau)
df <- mkt2(g1, g2, group_labels = c("Condition1", "Condition2"))

p <- ggplot(df, aes(x = obs)) + theme_classic() + 
  stat_density(aes(colour = gr), geom="line",position="identity", size=1) +
  scale_colour_viridis_d(end = 0.8) +
  coord_cartesian(xlim = c(0, 2500)) +
  theme(axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 16, colour="black"),
        axis.text.y = element_text(size = 16, colour="black"),
        axis.title.y = element_text(size = 18),
        legend.key.width = unit(1.5,"cm"),
        legend.position = c(0.55,0.75),
        legend.direction = "vertical",
        legend.text=element_text(size=16),
        legend.title=element_text(size=18),
        title = element_text(size=20)) +
  labs(x = "Reaction times", y = "Density", colour = "Groups") + 
  ggtitle("Spread difference")
p
p.spr.dist <- p
```

## Shift function
```{r}
out <- shifthd_pbci(df, nboot = 200, adj_ci = FALSE)
p <- plot_sf(out, plot_theme = 1)[[1]] + 
     theme(axis.text = element_text(size = 16, colour="black"))
p
p.spr.sf <- p
```

# Early difference

## KDE
```{r}
nt <- 1000
# ex Gaussian parameters
mu <- 500
sigma <- 50
tau <- 200
ES <- 1.5

set.seed(777)
g1 <- rexgauss(nt, mu = mu, sigma = sigma, tau = tau)
md.g1 <- median(g1)
q3.g1 <- quantile(g1, probs = 0.75)
# g1 <- (g1 - md.g1) * ES + md.g1 # spread around the median
g1 <- (g1 - q3.g1) * ES + q3.g1 # spread around the 3rd quartile
g2 <- rexgauss(nt, mu = mu, sigma = sigma, tau = tau)
df <- mkt2(g1, g2, group_labels = c("Condition1", "Condition2"))

p <- ggplot(df, aes(x = obs)) + theme_classic() + 
  stat_density(aes(colour = gr), geom="line",position="identity", size=1) +
  scale_colour_viridis_d(end = 0.8) +
  coord_cartesian(xlim = c(0, 2500)) +
  theme(axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 16, colour="black"),
        axis.text.y = element_text(size = 16, colour="black"),
        axis.title.y = element_text(size = 18),
        legend.key.width = unit(1.5,"cm"),
        legend.position = c(0.55,0.75),
        legend.direction = "vertical",
        legend.text=element_text(size=16),
        legend.title=element_text(size=18),
        title = element_text(size=20)) +
  labs(x = "Reaction times", y = "Density", colour = "Groups") +
  ggtitle("Early difference")
p
p.early.dist <- p
```

## Shift function
```{r}
out <- shifthd_pbci(df, nboot = 200, adj_ci = FALSE)
p <- plot_sf(out, plot_theme = 1)[[1]] + 
     theme(axis.text = element_text(size = 16, colour="black"))
p
p.early.sf <- p
```

# Late difference

## KDE
```{r}
nt <- 1000
# ex Gaussian parameters
mu <- 500
sigma <- 50
tau <- 200
ES <- 100

set.seed(777)
g1 <- rexgauss(nt, mu = mu, sigma = sigma, tau = tau + ES)
g2 <- rexgauss(nt, mu = mu, sigma = sigma, tau = tau)
df <- mkt2(g1, g2, group_labels = c("Condition1", "Condition2"))

p <- ggplot(df, aes(x = obs)) + theme_classic() + 
  stat_density(aes(colour = gr), geom="line",position="identity", size=1) +
  scale_colour_viridis_d(end = 0.8) +
  coord_cartesian(xlim = c(0, 2500)) +
  theme(axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 16, colour="black"),
        axis.text.y = element_text(size = 16, colour="black"),
        axis.title.y = element_text(size = 18),
        legend.key.width = unit(1.5,"cm"),
        legend.position = c(0.55,0.75),
        legend.direction = "vertical",
        legend.text=element_text(size=16),
        legend.title=element_text(size=18),
        title = element_text(size=20)) +
  labs(x = "Reaction times", y = "Density", colour = "Groups") + 
  ggtitle("Late difference")
p
p.tau.dist <- p
```

## Shift function
```{r}
out <- shifthd_pbci(df, nboot = 200, adj_ci = FALSE)
p <- plot_sf(out, plot_theme = 1)[[1]] + 
     theme(axis.text = element_text(size = 16, colour="black"))
p
p.tau.sf <- p
```



